---
title: 'Assignment-6: Problem 1'
output:
  html_document:
    df_print: paged
---

```{r}
#HARSH SHAH (001273963)
#DA5030 | 35664 | Intro Data Mining & Machine Learning | SEC 01 | Spring 2018
#Assignment: 6 
```


Dataset on student achievement
```{r}
#Import and read the student-mat dataset
student.mat <- read.csv("C:/Users/hshah/Desktop/DM-ML/Assignment 6 Regression-Linear, Multiple, Logistic/student/student-mat.csv", sep=";")

# To analyze the structure of the dataset
str(student.mat)

# To analyze the type of distribution of the underlying dataset
hist(student.mat$G3)
```

Problem-1: Question 1
Create scatter plots and pairwise correlations between four continuous variables and the final grade (G3) using the pairs.panels() function in R. Pick the variables you believe are most useful.
```{r}
df1 <- student.mat
df1

str(df1)

# Loaded "psych" library to use the following underlying functions for analysis of the correlations between the target and predictor variables.
library(psych)

pairs.panels(df1[c("G3", "studytime", "famrel", "goout", "absences")])
```

Problem-1: Question 2
Build a multiple regression model predicting final math grade (G3) using as many features as you like but you must use at least four. Include at least one categorical variables and be sure to properly convert it to dummy codes. Select the features that you believe are useful -- you do not have to include all features.
```{r}
df2 <- data.frame(student.mat)
str(df2)

# Creating dummy code variables in case of the categorical variables with more that 2 factors, in order to carry over the multiple regression model as per the requirement of the analysis. 
# Also used, only features that I believed are useful

guardian_dummy <- df2$guardian
guardian_mother <- ifelse(guardian_dummy == "mother", 1, 0)
guardian_other <- ifelse(guardian_dummy == "other", 1, 0)

paid_dummy <- df2$paid
paid_dummy <- ifelse(paid_dummy == "yes", 1, 0)

famsize_dummy <- df2$famsize
famsize_dummy <- ifelse(famsize_dummy == "GT3", 1, 0)

famsup_dummy <- df2$famsup
famsup_dummy <- ifelse(famsup_dummy == "yes", 1, 0)

str(df2)

df2 <- cbind(paid_dummy, df2)
df2 <- cbind(famsize_dummy, df2)
df2 <- cbind(famsup_dummy, df2)
df2 <- cbind(guardian_mother, df2)
df2 <- cbind(guardian_other, df2)


# Multiple regression model
b <- lm(G3 ~ traveltime+goout+absences+guardian_mother+guardian_other+paid_dummy+famsize_dummy+famsup_dummy, data = df2)

summary(b)
```

Problem-1: Question 3
Use stepwise backward elimination to remove all non-significant variables and then state the final model as an equation. State the backward elimination measure you applied (p-value, AIC, Adjusted R2).
```{r}
# Applied AIC method with backward elimination in order to remove all non-significant variables using "step()" function
step(b)

str(df2)

# Training a final multiple regression model with statistically significant feature variables
final_lm <- lm(G3 ~ traveltime+goout+paid_dummy+famsize_dummy+guardian_other, data = df2)

summary(final_lm)


# State the final model equation, in order to carry out regression of new cases with highest statistical significance

# 13.5165 - 0.7143(traveltime) - 0.5506(goout) + 0.8856(paid_dummy) - 0.9159(famsize_dummy) - 1.3523(guardian_other)
```

Problem-1 Question 4
Calculate the 95% confidence interval for a prediction -- you may choose any data you wish for some new student.
```{r}
pred.m  <- lm(formula = G3 ~ traveltime + goout + paid_dummy + famsize_dummy + guardian_other, data = df2)
summary(pred.m)

str(df2)
# New case with data of the student
new_d <- data.frame(traveltime=1, goout=3, paid_dummy=1, famsize_dummy=1, guardian_other=0)
new_d

# Predicting the result of the student with new student case data; here by passing parameter with interval = "confidence", it will apparantly carry out prediction with 95% confidence interval
v <- predict(pred.m, new_d, interval = "confidence")
v

# 95% confidence interval (10.36038, 11.8801) #fit = 11.12024
#v+CI_interval = 10.36038
#v-CI_interval = 11.8801
```

Problem-1: Question 5
What is the RMSE for this model -- use the entire data set for both training and validation.
```{r}
df2$predicted <- predict(pred.m)   # Save the predicted values
df2$residuals <- residuals(pred.m) # Save the residual values

# Quick look at the actual, predicted, and residual values
library(dplyr)
df2 %>% select(G3, predicted, residuals) %>% head()
RMSE <- sqrt(mean((df2$residuals)^2))
RMSE #RMSE of the model
```




